---
title: "portfolio 9"
output: html_document
date: "2023-04-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
df <- read_excel("/Users/lindsaystall/Downloads/PIRUS_Public_May2020.xlsx")
df[df == -99 | df == -88] <- NA
```

```{r}
ggplot(data = df, aes(x = Age)) +
  geom_histogram(binwidth = 5, color = "black", fill = "orchid4") +
  labs(title = "Distribution of Age", x = "Age", y = "Frequency")
```

The mode age is around 25.

```{r}
marital_counts <- data.frame(table(df$Marital_Status))

ggplot(marital_counts, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "pink2") +
  xlab("Marital Status") +
  ylab("Frequency") +
  ggtitle("Distribution of Marital Status")
```

1 = Single (never married)
2 = Married (religious or civil marriage qualifies) 
3 = Divorced or Separated
4 = Widowed

The terrorists in this dataset are very often single. I wonder why??????????


```{r}
gender_count <- data.frame(count = c(42, 194), gender = c("Female", "Male"))

ggplot(gender_count, aes(x = "", y = count, fill = gender)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c("mediumspringgreen", "hotpink1")) +
  theme_void() +
  labs(fill = "Gender")
```

The vast majority of the terrorists were men, but I'm honestly surprised to see that many female terrorists represented...

```{r}
percent_female <- mean(df$Gender == 1, na.rm = TRUE) * 100
cat(paste0("Percentage of females: ", round(percent_female, 1), "%\n"))
```

Around 10% of the terrorists were women.

```{r}
female_data <- subset(df, Gender == 1)

ggplot(female_data, aes(x = Radical_Signif_Other, fill = factor(Radical_Signif_Other))) +
  geom_bar() +
  xlab("Radical Significant Other") +
  ylab("Count") +
  ggtitle("Radical Significant Other for Female Terrorist") +
  scale_fill_discrete(name = "Radical Significant Other", 
                      labels = c("No or Single", "Legal Activities Only", 
                                 "Non-Violent Illegal Activities Only", 
                                 "Extremist Violence"))
```

Women who engaged in terrorist violence were very often in a relationship with another individual engaged in extremeist violence.


```{r}
male_data <- subset(df, Gender == 2)

ggplot(male_data, aes(x = Radical_Signif_Other, fill = factor(Radical_Signif_Other))) +
  geom_bar() +
  xlab("Radical Significant Other") +
  ylab("Count") +
  ggtitle("Radical Significant Other for Male Terrorists") +
  scale_fill_discrete(name = "Radical Significant Other", 
                      labels = c("No or Single", "Legal Activities Only", 
                                 "Non-Violent Illegal Activities Only", 
                                 "Extremist Violence"))
```

Male terrorists by an large were single or dating someone not engaged in terorism.

